---
title: "Model calibration"
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r libraries, echo = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(rstan)
library(purrr)
library(rethinking)
library(gridExtra)
library(bayesplot)
library(tidyr)
devtools::load_all("../../R_Packages/readsdr/")
```

```{r}
generate_recovery_df <- function(model_obj) {
  model_obj$model_df %>% select(time, Recovered) %>% 
  filter(time - trunc(time) == 0) %>% 
  mutate(recovery = Recovered - lag(Recovered, default = Recovered[1]),
         recovery = round(recovery)) %>% 
  select(time, recovery) %>% 
  filter(time != 0)
}
```


# 1. SYNTHETIC DATA

```{r SIR_incidence_data, fig.height = 2.5, cache = TRUE}
source('./deterministic_models/SIR.R')
source('./deterministic_models/SEIR.R')

SIR_output          <- SIR(initInfected = 10, effectiveContacts = 1, 
                           recoveryTime = 2)
SIR_df              <- SIR_output $incidence_data %>% mutate(model = "SIR")
synthetic_incidence <- SIR_output$incidence_data$incidence 

SEIR_output          <- SEIR(initInfected = 10, effectiveContacts = 1, 
                             latentPeriod = 1, recoveryTime = 2)
SEIR_df              <- SEIR_output$incidence_data %>% mutate(model = "SEIR")
synthetic_incidence2 <- SEIR_output$incidence_data$incidence

incidence_df <- bind_rows(SIR_df, SEIR_df)

ggplot(incidence_df, aes(x = time, y = incidence, group = model)) +
  geom_line(aes(colour = model)) +
  scale_colour_brewer(palette = "Greys") +
  geom_point(aes(shape = model)) +
  theme_test()
```

## 1.1 SIR

### 1.1.1 Fitting one parameter (Effective contacts)

\hfill\break

```{r th_oneParam_SIR_normal, cache = TRUE}
source("./stan_utils.R")

func_params <- list(recovery_delay = 0.5)
stan_fun    <- generate_stan_function("SIR", func_params)
stan_data   <- generate_data_block("normal")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("normal")
initStocks  <- list(Susceptible = 990, Infected = 10, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SIR", initStocks)
stan_model  <- generate_stan_model(likelihood = "normal",
                                  n_params = 1)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SIR.stan")

stan_fit <- fit_stan_model(
  stan_file = "SIR.stan", n_constants = 1, synthetic_incidence, warm_up = 2000, 
  iterations = 4000, seed = 191165034, n_difeq = 3)
```

```{r g_th_oneParam_SIR_normal, cache = TRUE}
posterior_df <- as.data.frame(stan_fit) %>% 
  rename(effective_contacts = 'params[1]')

credible_interval <- HPDI(posterior_df$effective_contacts, prob = 0.95)
mean_param        <- mean(posterior_df$effective_contacts)
median_param      <- median(posterior_df$effective_contacts)

hist.y <- density(posterior_df$effective_contacts) %$% 
  data.frame(x = x, y = y) %>% 
  mutate(area = x >= credible_interval[1] & x <= credible_interval[2])

g1 <- ggplot(hist.y, aes(x = x)) + 
  geom_line(aes(y = y)) +
  geom_text(aes(x = mean_param + 0.0014, y = 700), 
            label = paste0("mean   = ", round(mean_param, 4)),
            colour = "#1261A0", size = 3) +
  geom_text(aes(x = mean_param + 0.0014, y = 630), 
            label = paste0("median = ", round(median_param, 4)),
            colour = "#1261A0", size = 3) +
  annotate("text", x = mean_param + 0.0014, y = 550, 
           label = paste0("[",
                          round(credible_interval[1], 4),",",
                          round(credible_interval[2], 4), "]"),
           size = 2,
           colour = "grey") +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = area)) +
  scale_fill_manual(values = c(NA, "lightgrey")) + 
  geom_vline(aes(xintercept = mean_param),
            color = "#1261A0", linetype ="dashed", size = 1) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = "Effective contacts", title = "Normal likelihood")
```

```{r th_oneParam_SIR_poisson, cache = TRUE}
source("./stan_utils.R")
# y is now discrete
stan_data_poisson <- "
data {
  int<lower = 1> n_obs; // Number of weeks sampled
  int<lower = 1> n_params; // Number of model parameters
  int<lower = 1> n_difeq; // Number of differential equations in the system
  int y[n_obs]; 
  real t0; // Initial time point (zero)
  real ts[n_obs]; // Time points that were sampled
}
"

stan_parameters_poisson <- "
parameters {
  real<lower = 0> params[n_params]; // Model parameters
}
"

func_params   <- list(recovery_delay = 0.5)
stan_function <- generate_stan_function("SIR", func_params)
initStocks    <- list(Susceptible = 990, Infected = 10, Recovered = 0)
stan_td       <- generate_transformed_data_block()
stan_tp       <- generate_transformed_parameters("SIR", initStocks)
stan_model    <- generate_stan_model(likelihood = "poisson",
                                  n_params = 1)

stan_text <- paste(stan_function, stan_data_poisson, stan_td, 
                   stan_parameters_poisson, stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SIR_1_par.stan")

stan_fit <- fit_stan_model(
  stan_file = "SIR_1_par.stan", n_constants = 1, synthetic_incidence, 
  warm_up = 2000, iterations = 4000, seed = 466899087, n_difeq = 3)
```

```{r g_th_oneParam_SIR_poisson, cache = TRUE}
#Effective constants posterior
posterior_df <- as.data.frame(stan_fit) %>% 
  rename(effective_contacts = 'params[1]')

credible_interval <- HPDI(posterior_df$effective_contacts, prob = 0.95)
mean_param        <- mean(posterior_df$effective_contacts)
median_param      <- median(posterior_df$effective_contacts)

hist.y <- density(posterior_df$effective_contacts) %$% 
  data.frame(x = x, y = y) %>% 
  mutate(area = x >= credible_interval[1] & x <= credible_interval[2])

g2 <- ggplot(hist.y, aes(x = x)) + 
  geom_line(aes(y = y)) +
  geom_text(aes(x = mean_param + 0.026, y = 35), 
            label = paste0("mean   = ", round(mean_param, 4)),
            colour = "#1261A0", size = 3) +
  geom_text(aes(x = mean_param + 0.026, y = 32), 
            label = paste0("median = ", round(median_param, 4)),
            colour = "#1261A0", size = 3) +
  annotate("text", x = mean_param + 0.026, y = 28, 
           label = paste0("[",
                          round(credible_interval[1], 4),",",
                          round(credible_interval[2], 4), "]"),
           size = 2,
           colour = "grey") +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = area)) +
  scale_fill_manual(values = c(NA, "lightgrey")) + 
  geom_vline(aes(xintercept = mean_param),
            color = "#1261A0", linetype ="dashed", size = 1) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = "Effective contacts", title = "Poisson likelihood")
```


```{r g_th_oneParam_SIR, fig.height = 3}
grid.arrange(g1, g2, nrow = 1)
```

### 1.1.2 Fitting two parameters

#### 1.1.2.1 Effective contacts & recovery delay

##### 1.1.2.1.1 Normal likelihood

```{r th_SIR_normal_two_params, cache = TRUE}
source('./stan_utils.R')

stan_fun     <- generate_stan_function("SIR")
stan_data    <- generate_data_block("normal")
stan_td      <- generate_transformed_data_block()
stan_params  <- generate_parameters_block("normal")
initStocks   <- list(Susceptible = 990, Infected = 10, Recovered = 0)
stan_tp      <- generate_transformed_parameters("SIR", initStocks)
normal_model <- c("params[1] ~ uniform(0, 500)", "params[2] ~ uniform(0, 20)",
                "sigma ~ cauchy( 0 , 2 )", "y ~ normal(incidence, sigma)")
stan_model   <- generate_model_text(normal_model)

stan_text <- paste(stan_fun, stan_data, stan_td, stan_params,stan_tp, 
                   stan_model, sep = "\n")

create_stan_file(stan_text, "SIR.stan")
```


###### 1.1.2.1.1.1 Right model (SIR fitting SIR incidence)

\hfill\break

```{r th_SIR_normal_two_params_right, cache = TRUE}
sf_normal_right <- fit_stan_model(
  n_constants = 2, synthetic_incidence, warm_up = 2000, iterations = 4000,
  seed = 39843593, n_difeq = 3)
```

```{r p_th_SIR_normal_two_params_right, fig.width = 9, fig.height = 3}
g <- mcmc_pairs(sf_normal_right, pars = c("params[1]", "params[2]", "sigma"))
print(g)
```
\hfill\break

```{r g_th_SIR_normal_two_params_right, cache = TRUE, fig.height = 4}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "recoveryProportion", old = "params[2]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = sf_normal_right,
  pars = c("effective_contacts", "recoveryTime"),
  rename_pars = rename_pars,
  x_pos          = c(1.004, 2.023),
  y_pos          = c(200, 56),
  y_pos_median   = c(185, 52),
  y_pos_interval = c(165, 46))

g <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)
print(g)
```
\newpage

###### 1.1.2.1.1.2 Wrong model (SIR fitting SEIR incidence)

```{r th_SIR_normal_two_params_wrong, cache = TRUE}
source('./stan_utils.R')

stan_fit_normal_wrong <- fit_stan_model(
  n_constants = 2, synthetic_incidence2, warm_up = 2000, iterations = 4000,
  seed = 1407382414, n_difeq = 3)
pairs(stan_fit_normal_wrong, pars = c("params[1]", "params[2]", "sigma"))
```

```{r g_th_SIR_normal_two_params_wrong, cache = TRUE}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "recoveryProportion", old = "params[2]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fit_normal_wrong,
  pars = c("effective_contacts", "recoveryTime"),
  rename_pars = rename_pars,
  x_pos = c(0.535, 4.6), 
  y_pos = c(50, 2.6),
  y_pos_median = c(48, 2.4), 
  y_pos_interval = c(45, 2.2))

g <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)
print(g)
```

\newpage
##### 1.1.2.1.2 Poisson likelihood

```{r th_SIR_poisson_two_params, cache = TRUE}
source("./stan_utils.R")

stan_fun    <- generate_stan_function("SIR")
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson")
initStocks  <- list(Susceptible = 990, Infected = 10, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SIR", initStocks)
stan_model  <- generate_stan_model(likelihood = "poisson",
                                  n_params = 2)
stan_text <- paste(stan_fun, stan_data, stan_td, stan_params,
                    stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SIR_poisson.stan")
```

###### 1.1.2.1.2.1 Right model (SIR fitting SIR incidence)

```{r th_SIR_poisson_two_params_right, cache = TRUE}
source("./stan_utils.R")

stan_fit_poisson_right <- fit_stan_model(
  stan_file = 'SIR_poisson.stan', n_constants = 2, synthetic_incidence, 
  warm_up = 2000, iterations = 4000, seed = 245391088, n_difeq = 3)

pairs(stan_fit_poisson_right, pars = c("params[1]", "params[2]"))
```

```{r g_th_SIR_poisson_two_params_right, cache = TRUE}
source("./graphs.R")
library(rethinking)
# Parameters posterior
posterior_df <- as.data.frame(stan_fit_poisson_right) %>% 
  rename(effective_contacts = 'params[1]',
         recoveryProportion = 'params[2]') %>% 
  mutate(recoveryTime = 1 / recoveryProportion)

params_df<- posterior_df %>% 
  select(effective_contacts, recoveryTime)

credible_intervals <- apply(params_df, 2, HPDI, prob = 0.95) %>% t() %>% 
  as.data.frame() %>% 
  rename(lower_interval = "|0.95", upper_interval = "0.95|") %>% 
  mutate(param = rownames(.))

means   <- apply(params_df, 2, mean)
medians <- apply(params_df, 2, median)

stats_df <- data.frame(stringsAsFactors = FALSE,
                       param = names(means), 
                       mean_value = means,
                       median_value = medians,
                       lower_interval = credible_intervals$lower_interval,
                       upper_interval = credible_intervals$upper_interval) %>% 
  mutate(mean_label     = paste0("mean   = ", round(mean_value, 3)),
         median_label   = paste0("median = ", round(median_value, 3)),
         interval_label = paste0("[ ", round(lower_interval, 3), ", ", 
                                 round(upper_interval, 3), " ]"),
         x_pos          = c(1.09, 2.45),
         y_pos_mean     = c(11, 2.25),
         y_pos_median   = c(10, 2.10),
         y_pos_interval = c(8.6, 1.90))

densities <- apply(params_df, 2, density) %>% lapply(function(densityObj){
    data.frame(x = densityObj$x, y = densityObj$y)
  }) %>% bind_rows(.id = "param") %>% 
  inner_join(credible_intervals) %>% 
  mutate(area = x >= lower_interval & x <= upper_interval)

g <- draw_multiple_densities(densities, stats_df, text_size = 4)

print(g)

```

\newpage
###### 1.1.2.1.2.2 Wrong model (SIR fitting SEIR incidence)

```{r th_SIR_poisson_two_params_wrong, cache = TRUE}
source("./stan_utils.R")

stan_fit_poisson_wrong <- fit_stan_model(
  stan_file = 'SIR_poisson.stan', n_constants = 2, synthetic_incidence2, 
  warm_up = 2000, iterations = 4000, seed = 1926466221, n_difeq = 3)

pairs(stan_fit_poisson_wrong, pars = c("params[1]", "params[2]"))
```

```{r g_th_SIR_poisson_two_params_wrong, cache = TRUE}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "recoveryProportion", old = "params[2]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fit_poisson_wrong,
  pars = c("effective_contacts", "recoveryTime"),
  rename_pars = rename_pars,
  x_pos = c(0.56, 4.8), 
  y_pos = c(20, 0.8),
  y_pos_median = c(19, 0.75), 
  y_pos_interval = c(18, 0.7))

g <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)
print(g)
```

\newpage
#### 1.1.2.2 Effective contacts & S0

##### 1.1.2.2.1 Normal likelihood

```{r th_SIR_normal_two_params2}
source('./stan_utils.R')

func_params  <-list(recovery_delay = 0.5)
stan_fun     <-generate_stan_function("SIR", func_params)
stan_data    <- generate_data_block("normal")
stan_td      <- generate_transformed_data_block()
stan_params  <- generate_parameters_block("normal", stocks = "S0")
initStocks   <- list(Susceptible = "S0", Infected = 10, Recovered = 0)
stan_tp      <- generate_transformed_parameters("SIR", initStocks)
normal_model <- c("params[1] ~ uniform(0, 500)", "S0 ~ normal(0, 1000000)",
                "sigma ~ cauchy( 0 , 2 )", "y ~ normal(incidence, sigma)")
stan_model   <- generate_model_text(normal_model)

stan_text <- paste(stan_fun, stan_data, stan_td, stan_params,stan_tp, 
                   stan_model, sep = "\n")

create_stan_file(stan_text, "SIR.stan")
```

###### 1.1.2.2.1.1 Right model (SIR fitting SIR incidence)
```{r th_SIR_normal_two_params_right2, cache = TRUE, fig.height = 4}
stan_fit_normal_right2 <- fit_stan_model(
  n_constants = 1, synthetic_incidence, warm_up = 2000, iterations = 4000,
  seed = 1694212849, n_difeq = 3)

pairs(stan_fit_normal_right2, pars = c("params[1]", "S0", "sigma"))
```

```{r g_th_SIR_normal_two_params_right2, cache = TRUE, fig.height = 4}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"))

graph_inputs <- generate_graph_inputs(
  stan_fit_normal_right2, pars = c("effective_contacts", "S0"),
  rename_pars = rename_pars,
  x_pos = c(1.0005, 994), 
  y_pos = c(700, 0.3),
  y_pos_median = c(650, 0.28), 
  y_pos_interval = c(600, 0.26))

g <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)
print(g)
```

\newpage

###### 1.1.2.2.1.2 Wrong model (SIR fitting SEIR incidence)

```{r th_SIR_normal_two_params_wrong2, cache = TRUE}
stan_fit_normal_wrong2 <- fit_stan_model(
  n_constants = 1, synthetic_incidence2, warm_up = 2000, iterations = 4000,
  seed = 779238554, n_difeq = 3)

pairs(stan_fit_normal_wrong2, pars = c("params[1]", "S0", "sigma"))
```

```{r g_th_SIR_normal_two_params_wrong2, cache = TRUE}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"))

graph_inputs <- generate_graph_inputs(
  stan_fit_normal_wrong2, pars = c("effective_contacts", "S0"),
  rename_pars = rename_pars,
  x_pos = c(0.76, 1550), 
  y_pos = c(90, 0.012),
  y_pos_median = c(85, 0.011), 
  y_pos_interval = c(80, 0.010))

g <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)
print(g)
```

\newpage
##### 1.1.2.2.2 Poisson distribution

```{r th_SIR_poisson_two_params2}
source("./stan_utils.R")

func_params <-list(recovery_delay = 0.5)
stan_fun    <-generate_stan_function("SIR", func_params)
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson", stocks = "S0")
initStocks  <- list(Susceptible = "S0", Infected = 10, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SIR", initStocks)
stat_model  <- c("params[1] ~ uniform(0, 500)", "S0 ~ normal(0, 100000)",
                 "y ~ poisson(incidence)")
stan_model  <- generate_model_text(stat_model)
stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SIR_poisson.stan")
```

###### 1.1.2.2.2.1 Right model (SIR fitting SIR incidence)

```{r th_SIR_poisson_two_params_right2, cache = TRUE}
source("./stan_utils.R")

stan_fit_poisson_right2 <- fit_stan_model(
  stan_file = 'SIR_poisson.stan', n_constants = 1, synthetic_incidence, 
  warm_up = 2000, iterations = 4000, seed = 1682200822, n_difeq = 3)

pairs(stan_fit_poisson_right2, pars = c("params[1]", "S0"))
```

```{r g_th_SIR_poisson_two_params_right2, cache = TRUE}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"))

graph_inputs <- generate_graph_inputs(
  stan_fit_poisson_right2, pars = c("effective_contacts", "S0"),
  rename_pars = rename_pars,
  x_pos = c(1.02, 1060), 
  y_pos = c(30, 0.011),
  y_pos_median = c(28, 0.01), 
  y_pos_interval = c(26, 0.009))

g <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)
print(g)
```

\newpage

###### 1.1.2.2.2.2 Wrong model (SIR fitting SEIR incidence)

```{r th_SIR_poisson_two_params_wrong2, cache = TRUE}
source("./stan_utils.R")

sf_poisson_wrong2 <- fit_stan_model(
  stan_file = 'SIR_poisson.stan', n_constants = 1, synthetic_incidence2, 
  warm_up = 2000, iterations = 4000, seed = 591867144, n_difeq = 3)

pairs(sf_poisson_wrong2, pars = c("params[1]", "S0"))
```

```{r g_th_SIR_poisson_two_params_wrong2, cache = TRUE}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"))

graph_inputs <- generate_graph_inputs(
  sf_poisson_wrong2, pars = c("effective_contacts", "S0"),
  rename_pars = rename_pars,
  x_pos = c(0.767, 1490), 
  y_pos = c(54, 0.008),
  y_pos_median = c(51, 0.0075), 
  y_pos_interval = c(48, 0.007))

g <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)
print(g)
```

\newpage

#### 1.1.2.3 Calibration

```{r g_th_SIR_two_params, cache = TRUE}

produce_SIR_fit <- function(fit, label) {
  posterior_df <- as.data.frame(fit) %>%
    rename(effective_contacts = 'params[1]',
           recoveryProportion = 'params[2]') %>% 
    mutate(recoveryTime = 1 / recoveryProportion)
  
  SIR_obj <- SIR(
  initInfected = 10,
  effectiveContacts = mean(posterior_df$effective_contacts), 
  recoveryTime = mean(posterior_df$recoveryTime))
  
  SIR_obj$incidence_data %>% mutate(model = label)
}

labels   <- c("fit_normal_SIR", "fit_poisson_SIR")
fits     <- c(sf_normal_right, stan_fit_poisson_right)

fitted_df <- map2_df(fits, labels, produce_SIR_fit)
df        <- bind_rows(SIR_df, fitted_df)

g1 <- ggplot(df, aes(x = time, y = incidence, group = model)) +
  geom_line(aes(colour = model), alpha = 0.5) +
  scale_color_manual(values = c("darkgreen", "blue", "grey")) +
  scale_y_continuous(limits = c(0,100)) +
  theme_test() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 5)) 

labels   <- c("fit_normal_SIR", "fit_poisson_SIR")
fits     <- c(stan_fit_normal_wrong, stan_fit_poisson_wrong)

fitted_df <- map2_df(fits, labels, produce_SIR_fit)
df        <- bind_rows(SEIR_df, fitted_df)

g2 <- ggplot(df, aes(x = time, y = incidence, group = model)) +
  geom_line(aes(colour = model), alpha = 0.5) +
  scale_color_manual(values = c("darkgreen", "blue", "grey")) +
  scale_y_continuous(limits = c(0,100)) +
  theme_test() +
  theme(legend.position = "bottom", 
        legend.text = element_text(size = 5))

grid.arrange(g1, g2, nrow = 1, top = "Fitting effective contacts & recovery delay")
```

```{r g_th_SIR_two_params2, cache = TRUE}

produce_SIR_fit <- function(fit, label) {
  posterior_df <- as.data.frame(fit) %>%
    rename(effective_contacts = 'params[1]')
  
  SIR_obj <- SIR(
  initInfected = 10,
  initSusceptible = mean(posterior_df$S0),
  effectiveContacts = mean(posterior_df$effective_contacts), 
  recoveryTime = 2)
  
  SIR_obj$incidence_data %>% mutate(model = label)
}

labels   <- c("fit_normal_SIR", "fit_poisson_SIR")
fits     <- c(stan_fit_normal_right2, stan_fit_poisson_right2)

fitted_df <- map2_df(fits, labels, produce_SIR_fit)
df        <- bind_rows(SIR_df, fitted_df)

g1 <- ggplot(df, aes(x = time, y = incidence, group = model)) +
  geom_line(aes(colour = model), alpha = 0.5) +
  scale_color_manual(values = c("darkgreen", "blue", "grey")) +
  scale_y_continuous(limits = c(0,100)) +
  theme_test() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 5)) 
#-------------------------------------------------------------------------------

labels   <- c("fit_normal_SIR", "fit_poisson_SIR")
fits     <- c(stan_fit_normal_wrong2, sf_poisson_wrong2)

fitted_df <- map2_df(fits, labels, produce_SIR_fit)
df        <- bind_rows(SEIR_df, fitted_df)

g2 <- ggplot(df, aes(x = time, y = incidence, group = model)) +
  geom_line(aes(colour = model), alpha = 0.5) +
  scale_color_manual(values = c("darkgreen", "blue", "grey")) +
  scale_y_continuous(limits = c(0,100)) +
  theme_test() +
  theme(legend.position = "bottom", 
        legend.text = element_text(size = 5))

grid.arrange(g1, g2, nrow = 1, top = "Fitting effective contacts & S0")
```



### 1.1.3 Fitting three parameters (Effective contacts, recovery delay & S0)

```{r susceptible_priors, cache = TRUE}
sus_priors <- c(
  "S0 ~ normal(990, 1);",
  "S0 ~ normal(791, 10);",
  "S0 ~ normal(791, 250);")
```


#### 1.1.3.1 Fitting one dataset (Incidence data)

##### 1.1.3.1.1 Normal distribution

```{r th_SIR_normal_three_params, cache = TRUE}
source('./stan_utils.R')

# stan parameters normal init value. Adds S0
stan_pn_init_value <- "
parameters {
  real<lower = 0> params[n_params]; // Model parameters
  real<lower = 0> sigma;
  real<lower = 0> S0; // Initial number of susceptibles
}
"
# S0 is now an unknown value
stan_transformed_parameters <- "
transformed parameters{
  real y_hat[n_obs, n_difeq]; // Output from the ODE solver
  real y0[n_difeq]; // Initial conditions
  real incidence[n_obs];
    
  y0[1] = S0;
  y0[2] = 10;
  y0[3] = 0;
  
  y_hat = integrate_ode_rk45(SIR, y0, t0, ts, params, x_r, x_i);
  
  incidence[1] =  y_hat[1, 2] + y_hat[1, 3];
  
  for (i in 1:n_obs-1) {
    incidence[i + 1] = y_hat[i + 1, 2] + y_hat[i + 1, 3] - y_hat[i, 2] - y_hat[i, 3] + 0.00001; 
  }
  
}
"

stan_fun  <- generate_stan_function("SIR")
stan_data <- generate_data_block("normal")
stan_td   <- generate_transformed_data_block()

stan_fits_normal <- vector(mode = "list", length = length(sus_priors))

for(i in seq_len(length(sus_priors))) {
  
  susceptible_prior <- sus_priors[i]
  
  stan_model <- generate_stan_model(susceptible_prior, likelihood = "normal",
                                    n_params = 3)
  
  stan_text <- paste0(stan_fun, stan_data, stan_td, stan_pn_init_value, 
                      stan_transformed_parameters, stan_model)
  
  create_stan_file(stan_text, "SIR.stan")
  
  stan_fit <- fit_stan_model(n_constants = 2, synthetic_incidence,
                             warm_up = 2000, iterations = 4000, 
                             seed = 1000000)
  stan_fits_normal[i] <- stan_fit
}
```


###### 1.1.3.1.1.1 Prior: S0 ~ Normal(990, 1)

```{r g1_th_SIR_normal_three_params, fig.height = 4}
source('./graphs.R')

pairs(stan_fits_normal[[1]], pars = c("params[1]", "params[2]", "S0", "sigma"))
```

```{r tp_th_SIR_normal_three_params, fig.height = 4}
stan_trace(stan_fits_normal[[1]], pars = c("params[1]", "params[2]", "S0", "sigma"))
```


###### 1.1.3.1.1.2 Prior: S0 ~ Normal(791, 10)

```{r p2_th_SIR_normal_three_params}
pairs(stan_fits_normal[[2]], pars = c("params[1]", "params[2]", "S0", "sigma"))
```

```{r g2_th_SIR_normal_three_params}

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "recoveryProportion", old = "params[2]"))


graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fits_normal[[2]],
  pars = c("effective_contacts", "recoveryTime", "S0"),
  rename_pars = rename_pars,
  x_pos = c(0.83, 4.75, 830),
  y_pos = c(20, 1.2, 0.035),
  y_pos_median = c(19, 1.15, 0.0335),
  y_pos_interval = c(17, 1.05, 0.0305))


g <- draw_multiple_densities(graph_inputs$densities,
                             graph_inputs$stats_df,
                             text_size = 2)

g <- g + labs(title = "Prior: normal(791, 10)")
print(g)
```

\newpage
###### 1.1.3.1.1.3 Prior: S0 ~ Normal(791, 250)

```{r p3_th_SIR_normal_three_params}
pairs(stan_fits_normal[[3]], pars = c("params[1]", "params[2]", "S0", "sigma"))
```

```{r g3_th_SIR_normal_three_params}

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "recoveryProportion", old = "params[2]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fits_normal[[3]],
  pars = c("effective_contacts", "recoveryTime", "S0"),
  rename_pars = rename_pars,
  x_pos = c(1.04, 3,1020),
  y_pos = c(10, 1.2, 0.01),
  y_pos_median = c(9.6, 1.15, 0.0095),
  y_pos_interval = c(8.9, 1.08, 0.0087))


g <- draw_multiple_densities(graph_inputs$densities,
                             graph_inputs$stats_df,
                             text_size = 2)

g <- g + labs(title = "Prior: normal (791,250)")
print(g)
```

\newpage

##### 1.1.3.1.2 Poisson distribution

```{r th_SIR_poisson_three_params, cache = TRUE}
source('./stan_utils.R')

stan_fun    <- generate_stan_function("SIR")
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
# stan parameters Poisson init value. Adds S0
stan_pp_init_value <- "
parameters {
  real<lower = 0> params[n_params]; // Model parameters
  real<lower = 0> S0; // Initial number of susceptibles
}
"

initStocks  <- list(Susceptible = "S0", Infected = 10, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SIR", initStocks)

stan_fits_poisson <- vector(mode = "list", length = length(sus_priors))

for(i in seq_len(length(sus_priors))) {
  
  susceptible_prior <- sus_priors[i]
  
  stan_model <- generate_stan_model(susceptible_prior,
                                    likelihood = "poisson",
                                    n_params = 3)
  
  stan_text <- paste0(stan_fun, stan_data, stan_td , stan_pp_init_value, 
                      stan_tp, stan_model)
  
  create_stan_file(stan_text, "SIR.stan")
  
  stan_fit <- fit_stan_model(n_constants = 2, synthetic_incidence,
                             warm_up = 2000, 
                             iterations = 8000, 
                             seed = 1168103287,
                             n_difeq = 3)
  stan_fits_poisson[i] <- stan_fit
}
```

###### 1.1.3.1.2.1 Prior: S0 ~ Normal(990, 1)
```{r p1_th_SIR_poisson_three_params}
pairs(stan_fits_poisson[[1]], pars = c("params[1]", "params[2]", "S0"))
```


```{r g1_th_SIR_poisson_three_params}
# source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "recoveryProportion", old = "params[2]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fits_poisson[[1]],
  pars = c("effective_contacts", "recoveryTime", "S0"),
  rename_pars = rename_pars,
  x_pos = c(1.07, 2.4, 993),
  y_pos = c(10, 2.2, 0.35),
  y_pos_median = c(9.5, 2.1, 0.335),
  y_pos_interval = c(9, 2.0, 0.305))


g <- draw_multiple_densities(graph_inputs$densities,
                             graph_inputs$stats_df,
                             text_size = 2)

g <- g + labs(title = "Prior: normal (990,1)")
print(g)

```

###### 1.1.3.1.2.2 Prior: S0 ~ Normal(790, 10)

```{r p2_th_SIR_poisson_three_params}
  pairs(stan_fits_poisson[[2]], pars = c("params[1]", "params[2]", "S0"))
```


```{r g2_th_SIR_poisson_three_params}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "recoveryProportion", old = "params[2]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fits_poisson[[2]],
  pars = c("effective_contacts", "recoveryTime", "S0"),
  rename_pars = rename_pars,
  x_pos = c(0.87, 5.5, 820),
  y_pos = c(10, 0.6, 0.035),
  y_pos_median = c(9.5, 0.55, 0.0335),
  y_pos_interval = c(9, 0.5, 0.0305))


g <- draw_multiple_densities(graph_inputs$densities, 
                             graph_inputs$stats_df,
                             text_size = 2)

g <- g + labs(title = "Prior: normal (791, 10)")
print(g)

```

\newpage
###### 1.1.3.1.2.3 Prior: S0 ~ Normal(790, 250)

```{r p3_th_SIR_poisson_three_params}
pairs(stan_fits_poisson[[3]], pars = c("params[1]", "params[2]", "S0"))
```

```{r g3_th_SIR_poisson_three_params}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "recoveryProportion", old = "params[2]"))
graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fits_poisson[[3]],
  pars = c("effective_contacts", "recoveryTime", "S0"),
  rename_pars = rename_pars,
  x_pos = c(1.26, 3.5, 1280),
  y_pos = c(3, 0.8, 0.003),
  y_pos_median = c(2.8, 0.75, 0.0028),
  y_pos_interval = c(2.5, 0.68, 0.0026))


g <- draw_multiple_densities(graph_inputs$densities, 
                             graph_inputs$stats_df,
                             text_size = 2)

g <- g + labs(title = "Prior: normal(791, 250)")
print(g)

```

\newpage

##### 1.1.3.1.3 Calibration

```{r th_3_params_calibration}

get_sim_dfs <- function(fit, label) {
  posterior_df <- as.data.frame(fit) %>%
    rename(effective_contacts = 'params[1]',
           recoveryProportion = 'params[2]') %>%
    mutate(recoveryTime = 1 / recoveryProportion)

  SIR_obj <- SIR(
  initInfected = 10,
  effectiveContacts = mean(posterior_df$effective_contacts),
  recoveryTime = mean(posterior_df$recoveryTime),
  initSusceptible = mean(posterior_df$S0))
  
  recovery_df <- generate_recovery_df(SIR_obj) %>% 
    mutate(model = label)
  
  list(incidence_df = SIR_obj$incidence_data %>% mutate(model = label),
       recovery_df  = recovery_df)
}

sf_normal <- stan_fits_normal
sf_normal[[1]] <- NULL
sf_poisson <- stan_fits_poisson

labels_normal <- c("Normal (790,10)", "Normal (790, 250)")
labels_poisson <- c("Normal (990, 1)", labels_normal) 

sim_dfs_normal      <- map2(sf_normal, labels_normal, get_sim_dfs) 

incidence_df_normal <- map_df(sim_dfs_normal, "incidence_df") %>%
  rename(value = incidence) %>% 
  mutate(likelihood = "normal", var = "incidence")

R_data <- generate_recovery_df(SIR_output) %>% 
  mutate(model = "actual data", var = "recovery") %>% 
  rename(value = recovery)

recovery_df_normal  <- map_df(sim_dfs_normal, "recovery_df") %>%
  rename(value = recovery) %>% 
  mutate(likelihood = "normal", var = "recovery")

sim_dfs_poisson      <- map2(sf_poisson, labels_poisson, get_sim_dfs)

incidence_df_poisson <- map_df(sim_dfs_poisson , "incidence_df") %>%
  rename(value = incidence) %>%
  mutate(likelihood = "poisson", var = "incidence")

recovery_df_poisson  <- map_df(sim_dfs_poisson, "recovery_df") %>%
  rename(value = recovery) %>%
  mutate(likelihood = "poisson", var = "recovery")

comparison_df <- bind_rows(incidence_df_normal, recovery_df_normal,
                           incidence_df_poisson, recovery_df_poisson) %>% 
  mutate(model_likelihood = paste(model, likelihood, sep = "-"))

inc_data <- SIR_df %>% mutate(model = "actual data", var = "incidence") %>% 
  rename(value = incidence)

actual_data_df <- bind_rows(R_data, inc_data) %>% select(-model)


ggplot(comparison_df, aes(x = time, y = value)) +
  geom_line(aes(group = model_likelihood, colour = likelihood), alpha = 0.5) +
  # scale_colour_manual(values = c("#ca0020", "#f4a582", "#404040","#bababa")) +
  facet_grid(vars(var), vars(model)) +
  geom_line(data = actual_data_df, aes(x = time, y = value), colour = "grey") +
  theme_test()

```

\newpage
#### 1.1.3.2 Fitting two datasets (Incidence data & recovery data)

```{r incidence_recovery_graphs, cache = TRUE}
R_data <- generate_recovery_df(SIR_output)

df_incidence <- SIR_df %>% rename(value = incidence, var = model) %>% 
  mutate(var = "incidence")
df_recovery  <- R_data %>% rename(value = recovery) %>% 
  mutate(var = "recovery")

df_rates <- bind_rows(df_incidence, df_recovery)

ggplot(df_rates, aes(x = time, y = value)) +
  geom_line() +
  facet_wrap(~var, ncol = 1) +
  theme_test()
```


```{r th_SIR_poisson_three_params_two_datasets, cache = TRUE}
source('./stan_utils.R')
stan_fun    <- generate_stan_function("SIR")

stan_data   <- paste(
  "data {",
  "  int<lower = 1> n_obs; // Number of weeks sampled",
  "  int<lower = 1> n_params; // Number of model parameters",
  "  int<lower = 1> n_difeq; // Number of differential equations in the system",
  "  int y[n_obs];",
  "  int y2[n_obs];",
  "  real t0; // Initial time point (zero)",
  "  real ts[n_obs]; // Time points that were sampled",
  "}", sep = "\n")

stan_td     <- generate_transformed_data_block()
# stan parameters Poisson init value. Adds S0
stan_par    <- generate_parameters_block("poisson", "S0")

initStocks  <- list(Susceptible = "S0", Infected = 10, Recovered = 0)
stan_tp     <- paste(
  "transformed parameters {",
  "  real y_hat[n_obs, n_difeq]; // Output from the ODE solver",
  "  real y0[n_difeq]; // Initial conditions",
  "  real incidence[n_obs];",
  "  real recovery[n_obs];",
  "  y0[1] = S0;",
  "  y0[2] = 10;",
  "  y0[3] = 0;",
  "  y_hat = integrate_ode_rk45(SIR, y0, t0, ts, params, x_r, x_i);",
  "  incidence[1] =  y_hat[1, 2] + y_hat[1, 3] - y0[2] - y0[3];",
  "  recovery[1] = y_hat[1, 3] - y0[3];",
  "  for (i in 1:n_obs-1) {",
  "    incidence[i + 1] = y_hat[i + 1, 2] + y_hat[i + 1, 3] - y_hat[i, 2] - y_hat[i, 3] + 0.000001; ",
  "    recovery[i + 1] = y_hat[i + 1, 3] - y_hat[i, 3] + 0.000001;",
  "  }",
  "}", sep = "\n")

stan_fits_2_datasets <- lapply(sus_priors, function(prior) {
  model_text <- c(
  "params[1] ~ uniform(0, 500)",
  "params[2] ~ uniform(0, 20)",
  prior,
  "y ~ poisson(incidence)",
  "y2 ~ poisson(recovery)"
  )
  
  stan_model  <- generate_model_text(model_text)
  
  stan_text <- paste(stan_fun, stan_data, stan_td , stan_par,
                     stan_tp, stan_model, sep = "\n")
  
  create_stan_file(stan_text, "SIR_1.1.3.2.stan")
  
  stan_d <- list(n_obs    = length(synthetic_incidence), n_params = 2, 
                 n_difeq  = 3, # number of differential equations
                 y = synthetic_incidence, y2 = R_data$recovery, t0 = 0, 
                 ts = 1:length(synthetic_incidence))

  # Test / debug the model:
  test <- stan("SIR_1.1.3.2.stan", data = stan_d, chains = 1, iter = 10,
               verbose = FALSE, refresh = 0)
  
  
  # Fit and sample from the posterior
  stan_fit <- stan(fit = test, data = stan_d, chains = 3, warmup = 2000,
                   iter   = 4000, cores  = 3, seed = 1825157671)
  
})
```

```{r}
mcmc_pairs(stan_fits_2_datasets[[1]], pars = c("params[1]", "params[2]", "S0"))
```

```{r}
mcmc_pairs(stan_fits_2_datasets[[2]], pars = c("params[1]", "params[2]", "S0"))
```

```{r}
mcmc_pairs(stan_fits_2_datasets[[3]], pars = c("params[1]", "params[2]", "S0"))
```

#### 1.1.3.2.4 Calibration

```{r}
sim_dfs <- map2(stan_fits_2_datasets, sus_priors, function(fit, label) {
  posterior_df <- as.data.frame(fit) %>%
    rename(effective_contacts = 'params[1]',
           recoveryProportion = 'params[2]') %>% 
    mutate(recoveryTime = 1 / recoveryProportion)
  
  SIR_obj <- SIR(
  initInfected = 10,
  initSusceptible = mean(posterior_df$S0),
  effectiveContacts = mean(posterior_df$effective_contacts), 
  recoveryTime = mean(posterior_df$recoveryTime))
  
  recovery_df <- generate_recovery_df(SIR_obj) %>% 
    mutate(model = label)
  
  list(incidence_df = SIR_obj$incidence_data %>% mutate(model = label),
       recovery_df  = recovery_df)

})

actual_incidence <- map_df(sus_priors, function(prior) {
  SIR_df %>% mutate(model = prior, source = "actual data", 
                    variable = "Incidence") %>% 
    rename(value = incidence)
} )

inc_comparison_df <- map_df(sim_dfs, "incidence_df") %>% 
  mutate(variable = "Incidence", source = "simulated") %>% 
  rename(value = incidence) %>% bind_rows(actual_incidence)

actual_recovery  <- map_df(sus_priors, function(prior) {
  R_data %>% mutate(model = prior, source = "actual data",
                    variable = "Recovery") %>% 
  rename(value = recovery)
})

rec_comparison_df <- map_df(sim_dfs, "recovery_df") %>% 
  mutate(variable = "Recovery", source = "simulated") %>%
  rename(value = recovery) %>% bind_rows(actual_recovery)

comparison_df     <- bind_rows(inc_comparison_df, rec_comparison_df)

ggplot(comparison_df, aes(x = time, y = value)) +
  geom_line(aes(colour = source, group = source), alpha = 0.5) +
  facet_grid(vars(variable), vars(model)) +
  theme_test()
```


\newpage
## 1.2 SEIR

### 1.2.1 Fit one parameter (Contact rate)

```{r th_SEIR_one_param, cache = TRUE}
func_params <- list(recovery_delay = 0.5, latent_period = 1)
stan_fun    <- generate_stan_function("SEIR", func_params)
initStocks  <- list(Susceptible = 990, Exposed = 0, Infected = 10, Recovered = 0)
stan_td     <- generate_transformed_data_block()
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)
```


```{r th_SEIR_one_param_normal, cache = TRUE}

stan_data   <- generate_data_block("normal")
stan_params <- generate_parameters_block("normal")

normal_model_1_params_SEIR <- c(
  "params[1] ~ uniform(0, 500)",
  "sigma     ~ cauchy( 0 , 2 )",
  "y         ~ normal(incidence, sigma)"
)
stan_model  <- generate_model_text(normal_model_1_params_SEIR)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR_normal.stan")

stan_fit_normal <- NULL

stan_fit_normal <- fit_stan_model(
  stan_file = "SEIR_normal.stan", n_constants = 1, 
  incidence_data = synthetic_incidence2, warm_up = 2000, iterations = 4000,
  seed = 1010839870, chains = 3, n_difeq = 4)
```

```{r g_th_SEIR_one_param_normal}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fit_normal,
  pars = "effective_contacts",
  rename_pars = rename_pars,
  x_pos          = c(1.0025),
  y_pos          = c(500),
  y_pos_median   = c(450),
  y_pos_interval = c(400))

g1 <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)

g1 <- g1 + labs(title = "Normal")
```

```{r th_SEIR_one_param_poisson, cache = TRUE}
source('./stan_utils.R')

stan_data   <- generate_data_block("poisson")
stan_params <- generate_parameters_block("poisson")

poisson_model_1_param <- c(
  "params[1] ~ uniform(0, 500)",
  "y         ~ poisson(incidence)"
)

stan_model  <- generate_model_text(poisson_model_1_param)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR_poisson.stan")

stan_fit_poisson <- NULL
stan_fit_poisson <- fit_stan_model(
  stan_file = "SEIR_poisson.stan", n_constants = 1, 
  incidence_data = synthetic_incidence2, warm_up = 2000, iterations = 4000, 
  seed = 1600632916, chains = 3, n_difeq = 4)
```

```{r g_th_SEIR_one_param_poisson}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = stan_fit_poisson,
  pars = "effective_contacts",
  rename_pars = rename_pars,
  x_pos          = c(1.030),
  y_pos          = c(25),
  y_pos_median   = c(22),
  y_pos_interval = c(19))

g2 <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)

g2 <- g2 + labs(title = "Poisson")

```

```{r g_th_SEIR_one_param, fig.height = 3}
grid.arrange(g1, g2, nrow = 1)
```


### 1.2.2 Fit two parameters (Contact rate & latent period)


```{r th_SEIR_two_param, cache = TRUE}
source('./stan_utils.R')
func_params <- list(recovery_delay = 0.5)
stan_fun    <- generate_stan_function("SEIR", func_params)
initStocks  <- list(Susceptible = 990, Exposed = 0, Infected = 10, Recovered = 0)
stan_td     <- generate_transformed_data_block()
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)
```


#### 1.2.2.1 Normal distribution

```{r th_SEIR_two_param_normal, cache = TRUE}
source('./stan_utils.R')

stan_data   <- generate_data_block("normal")
stan_params <- generate_parameters_block("normal")

normal_model_2_params_SEIR <- c(
  "params[1] ~ uniform(0, 500)",
  "params[2] ~ uniform(0, 20)",
  "sigma     ~ cauchy( 0 , 2 )",
  "y         ~ normal(incidence, sigma)"
)
stan_model  <- generate_model_text(normal_model_2_params_SEIR)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR_normal.stan")

stan_fit_normal <- NULL

stan_fit_normal <- fit_stan_model(
  stan_file      = "SEIR_normal.stan",
  n_constants    = 2,
  incidence_data = synthetic_incidence2,
  warm_up        = 2000,
  iterations     = 4000,
  seed           = 2008167798,
  chains         = 3,
  n_difeq        = 4)
```

```{r, fig.height= 4}
mcmc_pairs(stan_fit_normal, pars = c("params[1]", "params[2]", "sigma"))
stan_trace(stan_fit_normal, pars = c("params[1]", "params[2]"))
```


#### 1.2.2.2 Poisson distribution

```{r th_SEIR_poisson_two_params}
source('./stan_utils.R')

func_params <- list(recovery_delay = 0.5)
stan_fun    <- generate_stan_function("SEIR", func_params)
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson")
initStocks  <- list(Susceptible = 990, Exposed = 0, Infected = 10, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)

poisson_model_2_params_SEIR <- c(
  "params[1] ~ uniform(0, 500)",
  "params[2] ~ uniform(0, 20)", 
  "y         ~ poisson(incidence)"
)
stan_model  <- generate_model_text(poisson_model_2_params_SEIR)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR_poisson1222.stan")
```


#### 1.2.2.2.1 Right model (SEIR fitting SEIR)

```{r th_SEIR_two_param_poisson, cache = TRUE}
source('./stan_utils.R')

sf_right <- NULL
sf_right <- fit_stan_model(
  stan_file = "SEIR_poisson1222.stan", n_constants = 2, 
  incidence_data = synthetic_incidence2, warm_up = 2000, iterations = 4000, 
  seed = 435686929, chains = 3, n_difeq = 4)
```

```{r, fig.height= 4}
mcmc_pairs(sf_right, pars = c("params[1]", "params[2]"))
```

```{r}
source('./graphs.R')

rename_pars <- list(list(new = "effective_contacts", old = "params[1]"),
                    list(new = "incubation_proportion", old = "params[2]"))

graph_inputs <- generate_graph_inputs(
  stan_fit = sf_right,
  pars = c("effective_contacts", "latent_period"),
  rename_pars = rename_pars,
  x_pos          = c(1.06, 1.25),
  y_pos          = c(8, 2.5),
  y_pos_median   = c(7.5, 2.4),
  y_pos_interval = c(7, 2.3))

g <- draw_multiple_densities(graph_inputs$densities, graph_inputs$stats_df,
                             text_size = 2)
print(g)
```


#### 1.2.2.2.2 Wrong model (SEIR fitting SIR)

```{r th_SEIR_poisson_two_params_wrong, cache = TRUE}
source('./stan_utils.R')

stan_fit_wrong <- fit_stan_model(
  stan_file = "SEIR_poisson1222.stan", n_constants = 2, synthetic_incidence,
  warm_up = 2000, iterations = 4000, seed = 28803987, n_difeq = 4)
```

```{r, fig.height= 4}
mcmc_pairs(stan_fit_wrong, np = nuts_params(stan_fit_wrong), 
           pars = c("params[1]", "params[2]"))
```


### 1.2.3 Fit three parameters (Contact rate, latent period & recovery time)

#### 1.2.3.1 Normal distribution

```{r th_SEIR_three_params_normal, cache=TRUE}
source('./stan_utils.R')

stan_fun    <- generate_stan_function("SEIR")
stan_data   <- generate_data_block("normal")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("normal")
initStocks  <- list(Susceptible = 990, Exposed = 0, Infected = 10,
                    Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)


create_input_for_model_block <- function(prior_incubation_proportion,
                                         prior_recovery_proportion) {
  
   c("params[1] ~ uniform(0, 500)", prior_incubation_proportion,
     prior_recovery_proportion, "sigma     ~ cauchy( 0 , 2 )",
     "y         ~ normal(incidence, sigma)")
}

priors_list <- list(
  list(label = "Flat prior",
       incubation_proportion = "params[2] ~ uniform(0, 20)",
       recovery_proportion = "params[3] ~ uniform(0, 20)"),
  list(label = "Regularising prior",
       incubation_proportion = "params[2] ~ normal(0, 10)",
       recovery_proportion = "params[3] ~ normal(0, 10)"),
  list(label = "Strong prior - incorrect",
       incubation_proportion = "params[2] ~ normal(0, 10)",
       recovery_proportion = "params[3] ~ normal(0.35, 0.01)"),
  list(label = "Strong prior - correct",
       incubation_proportion = "params[2] ~ normal(0, 10)",
       recovery_proportion = "params[3] ~ normal(0.50, 0.01)"))

stan_fits_normal <- lapply(priors_list, function(priors) {
  normal_model_3_params_SEIR <- create_input_for_model_block(
    priors$incubation_proportion, priors$recovery_proportion)
  
  plot_label  <<- priors$label
  stan_model  <- generate_model_text(normal_model_3_params_SEIR)

  stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

  create_stan_file(stan_text, "SEIR1231.stan")
  
  stan_fit_normal <- NULL
  stan_fit_normal <- fit_stan_model(
    stan_file = "SEIR1231.stan", n_constants = 3,
    incidence_data = synthetic_incidence2, warm_up = 1000, iterations = 2000,
    seed = 264842530, chains = 3, n_difeq = 4)
})
```

```{r, fig.height=4}

stan_fit <- stan_fits_normal[[1]]
np <- nuts_params(stan_fit)

mcmc_pairs(
  stan_fit, np = np,
  pars = c("params[1]", "params[2]", "params[3]", "sigma"),
  max_treedepth = with(np, -1 + max(Value[Parameter == "treedepth__"])),
  grid_args = list(top = "Flat prior"))

```

\hfill\break
```{r, fig.height=4}
stan_fit <- stan_fits_normal[[2]]
np <- nuts_params(stan_fit)
mcmc_pairs(
  stan_fit, np = np,
  pars = c("params[1]", "params[2]", "params[3]", "sigma"),
  max_treedepth = with(np, -1 + max(Value[Parameter == "treedepth__"])),
  grid_args = list(top = "Regularising prior" ))
```


```{r, fig.height=4}
stan_fit <- stan_fits_normal[[3]]
np <- nuts_params(stan_fit)
mcmc_pairs(
  stan_fit, np = np,
  pars = c("params[1]", "params[2]", "params[3]", "sigma"),
  max_treedepth = with(np, -1 + max(Value[Parameter == "treedepth__"])),
  grid_args = list(top = "Strong prior - incorrect"))
```
\hfill\break

```{r, fig.height=4}
stan_fit <- stan_fits_normal[[4]]
np <- nuts_params(stan_fit)
mcmc_pairs(
  stan_fit, np = np,
  pars = c("params[1]", "params[2]", "params[3]", "sigma"),
  max_treedepth = with(np, -1 + max(Value[Parameter == "treedepth__"])),
  grid_args = list(top = "Strong prior - correct" ))
```


#### Poisson distribution

```{r th_SEIR_three_params_poisson, cache = TRUE}
source('./stan_utils.R')

stan_fun    <- generate_stan_function("SEIR")
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson")
initStocks  <- list(Susceptible = 990, Exposed = 0, Infected = 10, 
                    Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)

poisson_model_3_params_SEIR <- c(
  "params[1] ~ uniform(0, 500)",
  "params[2] ~ normal(0, 10)",
  "params[3] ~ normal(0.5, 0.01)",
  "y         ~ poisson(incidence)"
)
stan_model  <- generate_model_text(poisson_model_3_params_SEIR)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR.stan")

stan_fit_poisson <- NULL
stan_fit_poisson <- fit_stan_model(
  stan_file = "SEIR.stan", n_constants = 3, 
  incidence_data = synthetic_incidence2, warm_up = 1000, iterations = 2000, 
  seed = 1789833883, chains = 3, n_difeq = 4)

pairs(stan_fit_poisson, pars = c("params[1]", "params[2]", "params[3]"))
```


# 2. REAL DATA


## 2.1 Flu data 57


```{r fludata, cache = TRUE}
source('./stan_utils.R')


flu_data <- read_csv("./data/flu1957.csv") %>% 
  mutate(total = Age.0.4 + Age.5.14 + Age.15.44 + Age.45.over)

population     <- 8000
I0             <- flu_data %>% slice(1) %>% pull(total)
S0             <- population - I0
flu_incidence  <- flu_data %>% slice(-1) %>% pull(total)

df1 <- flu_data %>% select(week, total) %>% slice(-1) %>% 
  mutate(type = "data",week = week - 1) %>% rename(incidence = total)

flu_data_long <-  pivot_longer(flu_data, -week, names_to = "variable", "value")
  

g <- ggplot(flu_data_long, aes(x = week, y = value)) +
  geom_line(colour = "lightgrey") +
  facet_wrap(~variable) +
  geom_point() +
  theme_test()

print(g)
```

## 2.2 SIR AGGREGATED
### One parameter (Contact rate = variable, recovery delay = 1.5 days)

```{r SIR_Flu_poisson_1_param, cache = TRUE}
source('./stan_utils.R')

recovery_proportion <- 7 / 1.5 # the inverse of 1.5 days in weeks
func_params <- list(recovery_delay = round(recovery_proportion, 10))
stan_fun    <- generate_stan_function("SIR", func_params)
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson")
initStocks  <- list(Susceptible = S0, Infected = I0, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SIR", initStocks)
stan_model  <- generate_stan_model(likelihood = "poisson",
                                   n_params = 1)

stan_text <- paste(stan_fun, stan_data, stan_td, stan_params,
                    stan_tp, stan_model, sep = "\n \n")

create_stan_file(stan_text, "SIR.stan")

stan_fit_poisson <- fit_stan_model(n_constants    = 1, 
                           incidence_data = flu_incidence,
                           warm_up        = 2000, 
                           iterations     = 4000, 
                           seed           = 778780507,
                           chains         = 3,
                           n_difeq        = 3)
```

```{r SIR_Flu_normal_1_param, cache = TRUE}
source('./stan_utils.R')

recovery_proportion <- 7 / 1.5
func_params <- list(recovery_delay = round(recovery_proportion, 10))
stan_fun    <- generate_stan_function("SIR", func_params)
stan_data   <- generate_data_block("normal")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("normal")
initStocks  <- list(Susceptible = S0, Infected = I0, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SIR", initStocks)
stan_model  <- generate_stan_model(likelihood = "normal",
                                  n_params = 1)

stan_text <- paste(stan_fun, stan_data, stan_td, stan_params,
                    stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SIR.stan")

stan_fit_normal <- fit_stan_model(n_constants    = 1,
                                  incidence_data = flu_incidence,
                                  warm_up        = 2000, 
                                  iterations     = 4000, 
                                  seed           = 1201008965,
                                  chains         = 3,
                                  n_difeq        = 3)

```



```{r g_SIR_Flu_1_param, cache=TRUE}
posterior_df <- as.data.frame(stan_fit_poisson) %>% 
  rename(effective_contacts = 'params[1]')

mean_ec <- mean(posterior_df$effective_contacts)

source('./deterministic_models/SIR.R')

model_output <- SIR(initInfected      = I0,
                    initSusceptible   = S0,
                    effectiveContacts = mean_ec, 
                    recoveryTime      = 1 / recovery_proportion,
                    start             = 0,
                    finish            = 18)

df2 <- model_output$incidence_data %>% rename(week = time) %>% 
  mutate(type = "Poisson")

posterior_df <- as.data.frame(stan_fit_normal) %>% 
  rename(effective_contacts = 'params[1]')

mean_ec <- mean(posterior_df$effective_contacts)

model_output <- SIR(initInfected      = I0,
                    initSusceptible   = S0,
                    effectiveContacts = mean_ec, 
                    recoveryTime      = 1 / recovery_proportion,
                    start             = 0,
                    finish            = 18)

df3 <- model_output$incidence_data %>% rename(week = time) %>% 
  mutate(type = "Gaussian")

df  <- bind_rows(df1, df2, df3)

ggplot(df, aes(x = week, y = incidence, group = type)) +
  geom_line(aes(colour = type, linetype = type)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  scale_colour_manual(values = c("lightgrey", "blue", "steelblue")) +
  theme_test()
```


### Two parameters (Contact rate & recovery delay are variables)

```{r real_SIR_Flu_poisson_2_param, cache = TRUE}
source('./stan_utils.R')

stan_fun    <- generate_stan_function("SIR")
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson")
initStocks  <- list(Susceptible = S0, Infected = I0, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SIR", initStocks)
stan_model  <- generate_stan_model(likelihood = "poisson",
                                  n_params = 2)
stan_text <- paste(stan_fun, stan_data, stan_td, stan_params,
                    stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SIR.stan")

stan_fit_poisson <- NULL
stan_fit_poisson <- fit_stan_model(
  n_constants    = 2, 
  incidence_data = flu_incidence,
  warm_up        = 2000, 
  iterations     = 4000, 
  seed           = 1410314945,
  chains         = 3)
```

```{r SIR_Flu_normal_2_param, cache = TRUE}
source('./stan_utils.R')

stan_fun     <- generate_stan_function("SIR")
stan_data    <- generate_data_block("normal")
stan_td      <- generate_transformed_data_block()
stan_params  <- generate_parameters_block("normal")
initStocks   <- list(Susceptible = S0, Infected = I0, Recovered = 0)
stan_tp      <- generate_transformed_parameters("SIR", initStocks)
normal_model <- c("params[1] ~ uniform(0, 500)", "params[2] ~ uniform(0, 20)",
                "sigma ~ cauchy( 0 , 2 )", "y ~ normal(incidence, sigma)")
stan_model   <- generate_model_text(normal_model)

stan_text <- paste(stan_fun, stan_data, stan_td, stan_params,
                    stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SIR.stan")

stan_fit_normal <- NULL
stan_fit_normal <- fit_stan_model(
  n_constants    = 2, 
  incidence_data = flu_incidence,
  warm_up        = 2000, 
  iterations     = 4000, 
  seed           = 1190756971,
  chains         = 3)

pairs(stan_fit_normal, pars = c("params[1]", "params[2]", "sigma"))
```

```{r graph_two_params, cache = TRUE}
fits   <- list(stan_fit_poisson, stan_fit_normal)
labels <- c("Poisson", "Gaussian") 

fits_df <- map2_df(fits, labels, function(fit, label) {
  
  samples <- as.data.frame(fit) %>% 
    rename(effective_contacts = 'params[1]',
         recovery_proportion = 'params[2]')
  
  mean_ec <- mean(samples$effective_contacts)
  mean_rd <- mean(samples$recovery_proportion)
  
  source('./deterministic_models/SIR.R')
  
  model_output <- SIR(initInfected      = I0,
                    initSusceptible     = S0,
                    effectiveContacts   = mean_ec, 
                    recoveryTime        = 1/ mean_rd,
                    start               = 0,
                    finish              = 18)
  
  model_output$incidence_data %>% rename(week = time) %>% 
  mutate(type = label)
})

df  <- bind_rows(df1, fits_df)

ggplot(df, aes(x = week, y = incidence, group = type)) +
  geom_line(aes(colour = type, linetype = type)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  scale_colour_manual(values = c("lightgrey", "blue", "steelblue")) +
  theme_test()
```

### Three parameters (Contact rate, recovery delay & S0)

```{r real_SIR_poisson_3_param, cache = TRUE}

stan_fun    <- generate_stan_function("SIR")
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
# stan parameters Poisson init value. Adds S0
stan_pp_init_value <- "
parameters {
  real<lower = 0> params[n_params]; // Model parameters
  real<lower = 0> S0; // Initial number of susceptibles
}
"
initStocks <- list(Susceptible = "S0", Exposed = 0, 
                    Infected = I0, Recovered = 0)
stan_tp    <- generate_transformed_parameters("SIR", initStocks)

S0_priors  <- c("S0 ~ normal(1914, 1000)", "S0 ~ normal(8000, 10)")

stan_fits_3_params <- purrr::map(S0_priors, function(S0_prior) {
  stat_model <- c("params[1] ~ uniform(0, 500)", "params[2] ~ uniform(0, 20)", 
                S0_prior, "y ~ poisson(incidence)")
  stan_model <- generate_model_text(stat_model)
  
  stan_text  <- paste0(stan_fun, stan_data, stan_td , stan_pp_init_value,
                       stan_tp, stan_model)
  
  create_stan_file(stan_text, "SIR_poisson.stan")
  
  stan_fit <- fit_stan_model(
  stan_file = "SIR_poisson.stan",n_constants = 2, flu_incidence, warm_up = 2000, 
    iterations = 4000, seed = 2131715700, n_difeq = 3)
  
  pairs(stan_fit, pars = c("params[1]", "params[2]", "S0"))
  stan_fit
})
```


```{r g_real_SIR_poisson_3_param, cache = TRUE}
RMSE <- function(s, a){
  sqrt(mean((s - a) ^ 2))
}

labels <- c("fit wrong S0", "fit right S0")

fits_df <- map2_df(stan_fits_3_params, labels, function(stan_fit, label) {
  
  samples <- as.data.frame(stan_fit) %>% 
    rename(effective_contacts = 'params[1]',
         recovery_proportion = 'params[2]')
  
  mean_ec <- mean(samples$effective_contacts)
  mean_rd <- mean(samples$recovery_proportion)
  mean_S0 <- mean(samples$S0)
  
  source('./deterministic_models/SIR.R')
  
  model_output <- SIR(
    initInfected      = I0,
    initSusceptible   = mean_S0,
    effectiveContacts = mean_ec, 
    recoveryTime      = 1 / mean_rd,
    start             = 0,
    finish            = 18)
  
    model_output$incidence_data %>% rename(week = time) %>%
    mutate(type = label)
})


df <- bind_rows(df1, fits_df)

errors <- map_dbl(labels, function(label) {
  simulated <- fits_df %>% filter(type == label) %>% pull(incidence)
  actual    <- df1$incidence
  RMSE(simulated, actual)
})

ggplot(df, aes(x = week, y = incidence, group = type)) +
  geom_line(aes(colour = type, linetype = type)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  scale_colour_manual(values = c("lightgrey", "blue", "steelblue")) +
  annotate("text", x = 8, y = 550, size = 3,
           label = paste0("RMSE = ",round(errors[[1]], 1)), colour = "steelblue") +
  annotate("text", x = 8, y = 400, size = 3,
           label = paste0("RMSE = ",round(errors[[2]], 1)), colour = "blue") +
  theme_test()
```




## 2.3 SEIR AGGREGATED

### 2.3.1 One parameter (Contact rate is variable, both latent period & recovery delay are 1 day)

```{r SEIR_Flu_poisson_1_param, cache = TRUE}
source('./stan_utils.R')

func_params <- list(latent_period  = round(7 / 1, 10), # inverse
                    recovery_delay = round(7 / 1, 10)) # inverse

stan_fun    <- generate_stan_function("SEIR", func_params)
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson")

initStocks  <- list(Susceptible = S0, Exposed = 0, Infected = I0, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)
stan_model  <- generate_stan_model(likelihood = "poisson", n_params = 1)
stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR.stan")

stan_fit_poisson <- NULL
stan_fit_poisson <- fit_stan_model(
  stan_file      = "SEIR.stan",
  n_constants    = 1,
  incidence_data = flu_incidence,
  warm_up        = 2000,
  iterations     = 4000,
  seed           = 609051934,
  chains         = 3,
  n_difeq        = 4)
```

```{r}
stan_trace(stan_fit_poisson, pars = "params[1]")
```


```{r SEIR_Flu_normal_1_param, cache=TRUE}
source('./stan_utils.R')

func_params <- list(latent_period  = round(7 / 1.5, 10), # inverse
                    recovery_delay = round(7 / 1.5, 10)) # inverse

stan_fun    <- generate_stan_function("SEIR", func_params)
stan_data   <- generate_data_block("normal")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("normal")
initStocks  <- list(Susceptible = S0, Exposed = 0, Infected = I0, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)
stan_model  <- generate_stan_model(likelihood = "normal", n_params = 1)
stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR.stan")

create_stan_file(stan_text, "SEIR.stan")
stan_fit_normal  <- NULL
stan_fit_normal  <- fit_stan_model(
  stan_file      = "SEIR.stan",
  n_constants    = 1,
  incidence_data = flu_incidence,
  warm_up        = 2000,
  iterations     = 4000,
  seed           = 563156067,
  chains         = 3,
  n_difeq        = 4)

pairs(stan_fit_normal, pars = c("params[1]", "sigma"))
```

```{r g_SEIR_Flu_1_param, cache = TRUE}
posterior_df <- as.data.frame(stan_fit_poisson) %>% 
  rename(effective_contacts = 'params[1]')

mean_ec <- mean(posterior_df$effective_contacts)


source('./deterministic_models/SEIR.R')

model_output <- SEIR(initInfected      = I0,
                    initSusceptible    = S0,
                    effectiveContacts  = mean_ec, 
                    recoveryTime       = 1.5 / 7,
                    latentPeriod       = 1.5 / 7,
                    start              = 0,
                    finish             = 18)

df2 <- model_output$incidence_data %>% rename(week = time) %>% 
  mutate(type = "Poisson")

df  <- bind_rows(df1, df2)

ggplot(df, aes(x = week, y = incidence, group = type)) +
  geom_line(aes(colour = type, linetype = type)) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_colour_manual(values = c("lightgrey", "blue")) +
  theme_test()
```

### Two parameters (effective contacts & latent period are variables, recovery delay is 1 day)

```{r real_SEIR_poisson_2_param, cache = TRUE}
source('./stan_utils.R')

func_params <- list(recovery_delay = round(7 / 1, 10))
stan_fun    <- generate_stan_function("SEIR", func_params)
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson")
initStocks  <- list(Susceptible = 7998, Exposed = 0, Infected = 2, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)

poisson_model_2_params_SEIR <- c(
  "params[1] ~ uniform(0, 500)",
  "params[2] ~ uniform(0, 200)", # the calibrated value moves as the interval is wider.
  "y         ~ poisson(incidence)"
)
stan_model  <- generate_model_text(poisson_model_2_params_SEIR)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR.stan")

stan_fit_poisson <- fit_stan_model(
  stan_file      = "SEIR.stan",
  n_constants    = 2,
  incidence_data = flu_incidence,
  warm_up        = 2000,
  iterations     = 4000,
  seed           = 1320624164,
  chains         = 3,
  n_difeq        = 4)

pairs(stan_fit_poisson, pars = c("params[1]", "params[2]"))
```

```{r real_SEIR_normal_2_param, cache=TRUE}
source('./stan_utils.R')
func_params <- list(recovery_delay = round(7 / 1, 10))
stan_fun    <- generate_stan_function("SEIR", func_params)
stan_data   <- generate_data_block("normal")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("normal")
initStocks  <- list(Susceptible = 7998, Exposed = 0, Infected = 2, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)

normal_model_2_params_SEIR <- c(
  "params[1] ~ uniform(0, 500)",
  "params[2] ~ uniform(0, 200)",
  "sigma     ~ cauchy( 0 , 2 )",
  "y         ~ normal(incidence, sigma)"
)
stan_model  <- generate_model_text(normal_model_2_params_SEIR)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR.stan")

stan_fit_normal <- NULL

stan_fit_normal <- fit_stan_model(
  stan_file      = "SEIR.stan",
  n_constants    = 2,
  incidence_data = flu_incidence,
  warm_up        = 2000,
  iterations     = 4000,
  seed           = 1471558969,
  chains         = 3,
  n_difeq        = 4)

pairs(stan_fit_normal, pars = c("params[1]", "params[2]"))
```


```{r SEIR_graph_two_params, cache=TRUE}

fits   <- list(stan_fit_poisson, stan_fit_normal)
labels <- c("Poisson", "Gaussian") 

fits_df <- map2_df(fits, labels, function(fit, label) {
  
  samples <- as.data.frame(fit) %>% 
    rename(effective_contacts = 'params[1]',
         incubation_proportion = 'params[2]')
  
  mean_ec <- mean(samples$effective_contacts)
  mean_ip <- mean(samples$incubation_proportion)
  
  source('./deterministic_models/SIR.R')
  
  model_output <- SEIR(initInfected      = I0,
                    initSusceptible    = S0,
                    effectiveContacts  = mean_ec, 
                    recoveryTime       = 1 / 7,
                    latentPeriod       = 1 / mean_ip,
                    start              = 0,
                    finish             = 18)
  
  model_output$incidence_data %>% rename(week = time) %>% 
  mutate(type = label)
})

df  <- bind_rows(df1, fits_df)

ggplot(df, aes(x = week, y = incidence, group = type)) +
  geom_line(aes(colour = type, linetype = type)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  scale_colour_manual(values = c("lightgrey", "blue", "steelblue")) +
  theme_test()
```


### Three parameters (effective contacts, latent period and recovery delay are variables)

### Poisson

```{r real_SEIR_poisson_3_param, cache=TRUE}
source('./stan_utils.R')
stan_fun    <- generate_stan_function("SEIR")
stan_data   <- generate_data_block("poisson")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("poisson")
initStocks  <- list(Susceptible = 7998, Exposed = 0, Infected = 2, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)

poisson_model_3_params_SEIR <- c(
  "params[1] ~ uniform(0, 500)",
  "params[2] ~ uniform(0, 50)",
  "params[3] ~ uniform(0, 50)",
  "y         ~ poisson(incidence)"
)
stan_model  <- generate_model_text(poisson_model_3_params_SEIR)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR.stan")

stan_fit_poisson <- fit_stan_model(
  stan_file      = "SEIR.stan",
  n_constants    = 3,
  incidence_data = flu_incidence,
  warm_up        = 2000,
  iterations     = 4000,
  seed           = 1320624164,
  chains         = 3,
  n_difeq        = 4)

pairs(stan_fit_poisson, pars = c("params[1]", "params[2]", "params[3]"))
```


### Normal
```{r real_SEIR_normal_3_param, cache=TRUE}
source('./stan_utils.R')

stan_fun    <- generate_stan_function("SEIR")
stan_data   <- generate_data_block("normal")
stan_td     <- generate_transformed_data_block()
stan_params <- generate_parameters_block("normal")
initStocks  <- list(Susceptible = 7998, Exposed = 0, Infected = 2, Recovered = 0)
stan_tp     <- generate_transformed_parameters("SEIR", initStocks)

normal_model_3_params_SEIR <- c(
  "params[1] ~ uniform(0, 500)",
  "params[2] ~ uniform(0, 50)",
  "params[3] ~ uniform(0, 50)",
  "sigma     ~ cauchy( 0 , 2 )",
  "y         ~ normal(incidence, sigma)"
)
stan_model  <- generate_model_text(normal_model_3_params_SEIR)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

create_stan_file(stan_text, "SEIR.stan")

stan_fit_normal <- NULL

stan_fit_normal <- fit_stan_model(
  stan_file      = "SEIR.stan",
  n_constants    = 3,
  incidence_data = flu_incidence,
  warm_up        = 1000,
  iterations     = 2000,
  seed           = 338192184,
  chains         = 3,
  n_difeq        = 4)

pairs(stan_fit_normal, pars = c("params[1]", "params[2]", "params[3]", "sigma"))
```


## 2.4 SEIR DISAGGREGATED

```{r}
source('./stan_utils.R')
stan_data  <- paste(
  "data {",
  "  int<lower = 1> n_obs; // Number of weeks sampled",
  "  int<lower = 1> n_params; // Number of model parameters",
  "  int<lower = 1> n_difeq; // Number of differential equations in the system",
  "  int y1[n_obs];",
  "  int y2[n_obs];",
  "  int y3[n_obs];",
  "  int y4[n_obs];",
  "  real t0; // Initial time point (zero)",
  "  real ts[n_obs]; // Time points that were sampled",
  "}", sep = "\n")

stan_td     <- generate_transformed_data_block()

stan_params <- generate_parameters_block("poisson")

model_file <- "./deterministic_models/4_cohorts_SEIR.stmx"
mdl <- read_xmile(model_file)
constant_names <- map_chr(mdl$description$constants, "name")
params <- constant_names[!constant_names%in% c("latent_period", 
                                               "recovery_time")]

counter <- 0

stan_init_stocks <- sapply(mdl$description$levels, function(level) {
  counter <<- counter + 1
  paste0("  y0[", counter, "] = ", level$initValue, ";")
}) %>% paste(collapse = "\n")

stan_tp <- paste(
  "transformed parameters{",
  "  real y_hat[n_obs, n_difeq]; // Output from the ODE solver",
  "  real y0[n_difeq]; // Initial conditions",
  "  real incidence1[n_obs];",
  "  real incidence2[n_obs];",
  "  real incidence3[n_obs];",
  "  real incidence4[n_obs];",
  stan_init_stocks,
  "  y_hat = integrate_ode_rk45(SEIR_extended, y0, t0, ts, params, x_r, x_i);",
  "  incidence1[1] =  y_hat[1, 3] + y_hat[1, 4] - y0[3] - y0[4];",
  "  incidence2[1] =  y_hat[1, 7] + y_hat[1, 8] - y0[7] - y0[8];",
  "  incidence3[1] =  y_hat[1, 11] + y_hat[1, 12] - y0[11] - y0[12];",
  "  incidence4[1] =  y_hat[1, 15] + y_hat[1, 16] - y0[15] - y0[16];",
  "  for (i in 1:n_obs-1) {",
  "    incidence1[i + 1] = y_hat[i + 1, 3] + y_hat[i + 1, 4] - y_hat[i, 3] - y_hat[i, 4] + 0.000001;",
  "    incidence2[i + 1] = y_hat[i + 1, 7] + y_hat[i + 1, 8] - y_hat[i, 7] - y_hat[i, 8] + 0.000001;",
  "    incidence3[i + 1] = y_hat[i + 1, 11] + y_hat[i + 1, 12] - y_hat[i, 11] - y_hat[i, 12] + 0.000001;",
  "    incidence4[i + 1] = y_hat[i + 1, 15] + y_hat[i + 1, 16] - y_hat[i, 15] - y_hat[i, 16] + 0.000001;",
  "   }",
  "}", sep = "\n")


```


### 2.4.1 Latent period & Recovery delay are equal to 1 day

```{r}
override_consts <- list(
  list(name = "recovery_time", value = round(1 / 7, 10)),
  list(name = "latent_period", value = round(1 / 7, 10)))

stan_fun   <- create_stan_function(model_file, "SEIR_extended", pars = params,
                                   override.consts = override_consts)

stan_model_text <- c(
"  params ~ normal(0, 0.0003)",
"  y1     ~ poisson(incidence1)",
"  y2     ~ poisson(incidence2)",
"  y3     ~ poisson(incidence3)",
"  y4     ~ poisson(incidence4)"
)
stan_model  <- generate_model_text(stan_model_text)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

filename <- "SEIR_241.stan"
create_stan_file(stan_text, filename)

cat(stan_text)
```


```{r seir_disaggregated241, cache = TRUE}
stan_d = list(n_obs    = length(flu_data$Age.0.4),
              n_params = 10, 
              n_difeq  = 16, # number of differential equations
              y1       = flu_data$Age.0.4,
              y2       = flu_data$Age.5.14,
              y3       = flu_data$Age.15.44,
              y4       = flu_data$Age.45.over,
              t0       = 0,
              ts       = 1:length(flu_data$Age.0.4))

# Test / debug the model:
test <- stan(filename, data = stan_d, chains = 1, iter = 10,
               verbose = FALSE, refresh = 0)

# Fit and sample from the posterior
stan_fit_2.4 <- stan(fit = test, data = stan_d, chains = 3, warmup = 1000,
                 iter   = 2000, cores  = 3, seed = 1099652217, refresh = 0)
```


```{r g_comparison_SEIR_dis_1_day}
posterior_df <- as.data.frame(stan_fit_2.4)

mean_incidences <- posterior_df %>% select(contains("incidence")) %>% 
  colMeans()

names_mi <- names(mean_incidences)

age_groups <- c("Age.0.4", "Age.5.14", "Age.15.44", "Age.45.over")

sim_data <- map_df(age_groups, function(ag) {
  i <- which(ag == age_groups)
  regex <- stringr::regex(paste0("incidence", i, "\\["))
  data.frame(week = 1:19, value = mean_incidences[grepl(regex, names_mi)], 
                            variable = ag)
}) %>% mutate(source = "sim data")

real_data <- flu_data_long %>% filter(variable != "total") %>% 
  mutate(source = "real data")

comparison_data <- bind_rows(sim_data, real_data)

ggplot(comparison_data, aes(x = week, y = value)) +
  geom_line(aes(group = source, colour = source)) +
  scale_colour_manual(values = c("lightgrey", "blue")) +
  facet_wrap(~variable) +
  theme_test()
```

```{r}
age_groups <- unique(comparison_data$variable)
map_dbl(age_groups, function(ag) {
  ag_data <- comparison_data %>% filter(variable == ag) %>% arrange(week)
  ag_sim  <- ag_data %>% filter(source == "sim data") %>% pull(value)
  ag_act  <- ag_data %>% filter(source == "real data") %>% pull(value)
  RMSE(ag_sim, ag_act)
})
```

### 2.4.2 Latent period & Recovery delay are equal to 1.5 days

```{r seir_disaggregated242, cache = TRUE}
override_consts <- list(
  list(name = "recovery_time", value = round(1.5 / 7, 10)),
  list(name = "latent_period", value = round(1.5 / 7, 10)))

stan_fun   <- create_stan_function(model_file, "SEIR_extended", pars = params,
                                   override.consts = override_consts)

stan_model_text <- c(
"  params ~ normal(0, 0.000600)",
"  y1     ~ poisson(incidence1)",
"  y2     ~ poisson(incidence2)",
"  y3     ~ poisson(incidence3)",
"  y4     ~ poisson(incidence4)"
)
stan_model  <- generate_model_text(stan_model_text)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

filename <- "SEIR_242.stan"
create_stan_file(stan_text, filename)

stan_d = list(n_obs    = length(flu_data$Age.0.4),
              n_params = 10, 
              n_difeq  = 16, # number of differential equations
              y1       = flu_data$Age.0.4,
              y2       = flu_data$Age.5.14,
              y3       = flu_data$Age.15.44,
              y4       = flu_data$Age.45.over,
              t0       = 0,
              ts       = 1:length(flu_data$Age.0.4))

# Test / debug the model:
test <- stan(filename, data = stan_d, chains = 1, iter = 10,
               verbose = FALSE, refresh = 0)

# Fit and sample from the posterior
stan_fit_2.4.2 <- stan(fit = test, data = stan_d, chains = 3, warmup = 1000,
                 iter   = 2000, cores  = 3, seed = 1742552100, refresh = 0)
```

```{r g_comparison_SEIR_dis_1_5_days}
posterior_df <- as.data.frame(stan_fit_2.4.2)

mean_incidences <- posterior_df %>% select(contains("incidence")) %>% 
  colMeans()

names_mi <- names(mean_incidences)

age_groups <- c("Age.0.4", "Age.5.14", "Age.15.44", "Age.45.over")

sim_data <- map_df(age_groups, function(ag) {
  i <- which(ag == age_groups)
  regex <- stringr::regex(paste0("incidence", i, "\\["))
  data.frame(week = 1:19, value = mean_incidences[grepl(regex, names_mi)], 
                            variable = ag)
}) %>% mutate(source = "sim data")

real_data <- flu_data_long %>% filter(variable != "total") %>% 
  mutate(source = "real data")

comparison_data <- bind_rows(sim_data, real_data)

ggplot(comparison_data, aes(x = week, y = value)) +
  geom_line(aes(group = source, colour = source)) +
  scale_colour_manual(values = c("lightgrey", "blue")) +
  facet_wrap(~variable) +
  theme_test()
```

```{r}
age_groups <- unique(comparison_data$variable)
map_dbl(age_groups, function(ag) {
  ag_data <- comparison_data %>% filter(variable == ag) %>% arrange(week)
  ag_sim  <- ag_data %>% filter(source == "sim data") %>% pull(value)
  ag_act  <- ag_data %>% filter(source == "real data") %>% pull(value)
  RMSE(ag_sim, ag_act)
})
```

### 2.4.3 Latent period & Recovery delay are equal to 2 days

```{r seir_disaggregated243, cache = TRUE}
override_consts <- list(
  list(name = "recovery_time", value = round(2 / 7, 10)),
  list(name = "latent_period", value = round(2 / 7, 10)))

stan_fun   <- create_stan_function(model_file, "SEIR_extended", pars = params,
                                   override.consts = override_consts)

stan_model_text <- c(
"  params ~ normal(0, 0.000040)",
"  y1     ~ poisson(incidence1)",
"  y2     ~ poisson(incidence2)",
"  y3     ~ poisson(incidence3)",
"  y4     ~ poisson(incidence4)"
)
stan_model  <- generate_model_text(stan_model_text)

stan_text   <- paste(stan_fun, stan_data, stan_td, stan_params,
                     stan_tp, stan_model, sep = "\n")

filename <- "SEIR_243.stan"
create_stan_file(stan_text, filename)

stan_d = list(n_obs    = length(flu_data$Age.0.4),
              n_params = 10, 
              n_difeq  = 16, # number of differential equations
              y1       = flu_data$Age.0.4,
              y2       = flu_data$Age.5.14,
              y3       = flu_data$Age.15.44,
              y4       = flu_data$Age.45.over,
              t0       = 0,
              ts       = 1:length(flu_data$Age.0.4))

# Test / debug the model:
test <- stan(filename, data = stan_d, chains = 1, iter = 10,
               verbose = FALSE, refresh = 0)

# Fit and sample from the posterior
stan_fit_2.4.3 <- stan(fit = test, data = stan_d, chains = 3, warmup = 1000,
                 iter   = 2000, cores  = 3, seed = 1250266639, refresh = 5)
```

```{r g_comparison_SEIR_dis_2_days}
posterior_df <- as.data.frame(stan_fit_2.4.3)

mean_incidences <- posterior_df %>% select(contains("incidence")) %>% 
  colMeans()

names_mi <- names(mean_incidences)

age_groups <- c("Age.0.4", "Age.5.14", "Age.15.44", "Age.45.over")

sim_data <- map_df(age_groups, function(ag) {
  i <- which(ag == age_groups)
  regex <- stringr::regex(paste0("incidence", i, "\\["))
  data.frame(week = 1:19, value = mean_incidences[grepl(regex, names_mi)], 
                            variable = ag)
}) %>% mutate(source = "sim data")

real_data <- flu_data_long %>% filter(variable != "total") %>% 
  mutate(source = "real data")

comparison_data <- bind_rows(sim_data, real_data)

ggplot(comparison_data, aes(x = week, y = value)) +
  geom_line(aes(group = source, colour = source)) +
  scale_colour_manual(values = c("lightgrey", "blue")) +
  facet_wrap(~variable) +
  theme_test()
```


```{r}
age_groups <- unique(comparison_data$variable)
map_dbl(age_groups, function(ag) {
  ag_data <- comparison_data %>% filter(variable == ag) %>% arrange(week)
  ag_sim  <- ag_data %>% filter(source == "sim data") %>% pull(value)
  ag_act  <- ag_data %>% filter(source == "real data") %>% pull(value)
  RMSE(ag_sim, ag_act)
})
```